Aquifer depletion exacerbates agricultural drought losses in the US High Plains

### 2023-08-30 Writing/NatureWater/1stRoundRevision/replies/reply_to_reviewers_R1.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{url}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\section{Reply to reviewer \#1}

This manuscript assesses the impact of aquifer thickness on agricultural
production during drought in the area overlying the High Plains aquifer
in the central United States. The authors use crop, meteorological, and
aquifer thickness data in regressions to develop relationships among
these quantities. Their most important finding is the impact of aquifer
thickness on productivity during drought and the ramifications for
aquifer conservation efforts. I have read this manuscript over a number
of times and found it of considerable interest and worthy of publication
in Nature Water. However, I do have a few concerns that I wanted to
bring to the attention of the authors. My most important concerns are as
follows (not necessarily in order of importance):

\textcolor{blue}{Thank you for the detailed and constructive feedback on our manuscript. We have endeavored to address all your comments and suggestions, and our replies to your comments can found in blue below (original comments in black)}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Aquifer thickness versus saturated thickness -- The authors speak of
  ``aquifer saturated thickness'', which is a redundant term. Remember, the definition of an aquifer is a saturated unit – there is no unsaturated aquifer. Please use ``aquifer
  thickness'', to avoid propagating this confusing usage into the
  future.
\end{enumerate}

\textcolor{blue}{We agree that there is some redundant use of words here. All the instances of ``saturated thickness'' and ``aquifer saturated thickness'' have been replaced by ``aquifer thickness'' in the revised manuscript.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Dependence on aquifer thickness -- I had a number of questions about
  the relationships displayed in Figure 1. First, the regression appears
  to have been done year by year. Are the quantiles recalculated each
  year to reflect that there may be changes between categories as time
  goes on? Second, when the authors state the lowest quantile goes from
  0 to 33\% are they assuming that the minimum thickness of 9 m is the
  zero point? Third, I assume that the county and year fixed effects are
  calculated as part of the regression. Are there any insights that can
  be gained from the interpretation of those quantities? Fourth, the
  authors need to point out that the same aquifer thickness can result
  in dramatically different transmissive characteristics. Thus, a
  relatively thin aquifer interval composed of coarse sands and gravels
  may be more resilient to drought than a much thicker aquifer dominated
  by fine sands and silts. Could that be a major player in the error
  bars in the plots? Please mention this at some point in the
  manuscript.
\end{enumerate}

\begin{itemize}
\item
  First point:
  \textcolor{blue}{Quantiles are not recalculated each year. If we do so, then the same quantile would represent different aquifer thickness range every year. When modeling the impact of aquifer thickness, this is not desirable because it would be expected that different ranges of aquifer thickness would have the same impact on a farmer's ability to buffer drought risk irrespective of year.}
\item
  Second point:
  \textcolor{blue}{Thank you for highlighting this point. For the 0\% quantile, we assume a aquifer thickness of 9m. This minimum of aquifer thickness was selected based on existing literature (Fenichel et al., 2016; Haacker et al., 2016; Deines et al., 2020), which typically defines 9m as the minimum aquifer thickness needed to provide a functioning water supply. We have modified the text on page 13 lines 298-302 of the revised manuscript to note these points.}
\item
  Third point:
  \textcolor{blue}{Yes, county and year fixed effects are calculated as part of the regressions. However, these variables can be interpreted simply as shifters that tell us how much higher yields in a single county are on average relative to another county and to account for trends in yields over time. As these variables do not interact with aquifer thickness in our analysis, there would be little insights gained from analyzing these variables - and indeed it is standard practice to not report and directly analyze fixed effects terms in econometric regressions used in statistical crop-climate modeling.}
\item
  Fourth point:
  \textcolor{blue}{This is an important point. You are correct that unobserved factors, such as aquifer heterogeneity influencing well yield responses to changes in aquifer thickness, are a likely underlying cause of the uncertainty ranges found in our analyses. Variations in aquifer conductivity, for example, would alter the relationship between saturated thickness and well capacity, thereby adding noise to the relationship between aquifer thickness and drought risk exposure. This noise is referred to as measurement error in econometric studies, and can result in attenuation bias. Specifically, measurement error in an explanatory variable (here, aquifer thickness) tends to bias your estimation in a way that makes the impact of the explanatory variable with measurement errors appear "less" influential than it truly is. So, in our context, our estimated impact of aquifer thickness on the share of irrigation is conservative because measurement error will be attenuating some of the true impact of aquifer thickness on drought risk exposure. Furthermore, our analysis shows that - even in the presence of these uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining aquifer thickness. This further reinforces the importance of aquifer conditions as a determinant of adaptive capacity to drought. To highlight these points, we have added the following text in the Discussion section on page 10 lines 213-227 of our revised manuscript:}
\end{itemize}

\textcolor{blue}{"Unobserved variations in aquifer properties will also introduce noise into the relationships between aquifer thickness, well yields, and drought risk. For example, a thinner aquifer interval composed of coarse sands and gravels may be able to supply comparable or higher well yields than a thicker aquifer dominated by fine sands and silt deposits, and hence offer greater resilience to water deficits despite similar or larger aquifer thickness [Butler et al., 2013, Korus and Hensen, 2020]. These measurement errors in the relationship between aquifer thickness and well yield will lead to attenuation bias [Bound and Krueger, 1991, Hyslop and Imbens, 2001], suggesting that our analysis is likely to be an underestimate of the true impacts of groundwater depletion on drought risk. Nonetheless, our analysis shows that - even in the presence of these measurement uncertainties - we are still able to identify statistically significant increases in drought risk exposure for irrigated agriculture as a function of declining aquiefr thickness. Improved data on spatial variations in aquifer properties and greater monitoring of real-world changes in well capacities alongside traditional water level measurements would help to further refine understanding of patterns of drought vulnerability across the HPA. This would enable more precise estimates to be made of critical tipping points in the resilience of irrigated agriculture to drought, which are needed to effectively design and target groundwater conservation policies."}

\textcolor{blue}{New references added:}

\begin{itemize}
\item
  \textcolor{blue}{Butler, J. J., Stotler, R. L., Whittemore, D. O., and Reboulet, E. C. (2013). Interpretation of water level changes in the High Plains aquifer in western Kansas. Groundwater, 51(2), 180-190.}
\item
  \textcolor{blue}{Korus, J. T., and Hensen, H. J. (2020). Depletion percentage and nonlinear transmissivity as design criteria for groundwater-level observation networks. Environmental Earth Sciences, 79(16), 382.}
\item
  \textcolor{blue}{John Bound and Alan B Krueger. The extent of measurement error in longitudinal earnings data: Do two wrongs make a right? Journal of labor economics, 9(1):1–24, 1991.}
\item
  \textcolor{blue}{Dean R Hyslop and Guido W Imbens. Bias from classical and other forms of measurement error. Journal of Business \& Economic Statistics, 19(4):475–481, 2001}
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Conditioned on drought risk exposure -- I wasn't sure what the authors
  meant by this in the second paragraph of Section 2.2. This must be
  related to the water deficit but the lack of explanation in the main
  text or the methods section left me wondering. Please clarify at some
  point in the manuscript.
\end{enumerate}

\textcolor{blue}{By "conditioned on drought risk exposure", we were referring to the fact that Figure 4 displays the relationship between aquifer thickness and the share of production area that is irrigated while also accounting for variations in climate across the aquifer. I.e. that the Figure shows the impact of aquifer thickness on share of production irrigated after removing the effects of climate on share of production irrigated. To improve clarity, we have modified this phrase to now state "...after accounting for spatio-temporal variations in weather conditions and drought risk exposure across the aquifer".}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Equation 2 -- Please define all terms in the equation and provide an
  explanation of what is meant by ``estimated semi-parametrically'\,'.
\end{enumerate}

\textcolor{blue}{We now added a more detailed summary of the mathematical expressions of the model with all the terms explained in Section 4.2 of our revised manuscript. The word semi-parametrically refers to the fact that generalized additive models, such as used in our analysis, estimate the impact of a variable using splines without functional form assumption. These are sometimes referred to as semi-parametric regression methods. However, we agree that the term is potentially confusing, and therefore have removed the term "semi-parametrically" in our revised manuscript.}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Kansas and California efforts -- I've seen a number of papers
  describing efforts in Kansas and California to address aquifer
  depletion via LEMAs (Kansas) and SGMA (California). The 1985-2016
  period examined by the authors was before most of the efforts in
  Kansas got started. My understanding is that the first LEMA was
  established in 2013 but then no more were set up until after 2016.
  This manuscript would be more impactful if the authors could speculate
  on how their relationships might change with greater attention to
  groundwater conservation. For example, I would think the decrease in
  yield the authors observe with negative water deficits would disappear
  when the irrigators are using soil moisture sensors and are focused on
  water-use efficiency. I would expect changes as well for tipping
  points.
\end{enumerate}

\textcolor{blue}{This is an interesting idea - thank you for raising it. We agree that efforts to improve aquifer management could help to mitigate some of the changes in productivity driven by climate and aquifer conditions in our analysis. Greater attention to groundwater conservation, and use of improved scheduling technologies such as soil moisture probes, would be expected to mitigate production losses in years with negative water deficits (i.e. wetter years) which we attribute to increased possibility for over-irrigation when well capacities are high (and hence when farmers do not face or observe physical scarcity). In years with large positive water deficits (i.e. drought years), improved irrigation water management would also be expected to mitigate some of the additional production risk caused by smaller aquifer thickness. However, we would still expect smaller aquifer thickness to result in elevated exposure to drought risk even in these circumstances, as declining well yields would create a physical limit to water supply which would stop farmers being able to fully meet crop water needs in these years. Formally testing these effects is beyond the scope and data of our study, but we agree that these are valuable areas for future work and have added the following text to the discussion of our revised manuscript on page 11 lines 233-246:}

\textcolor{blue}{"Improvements to irrigation scheduling practices and technologies, such as have been implemented as part of the SD6-LEMA (Local Enhanced Management Area) in Kansas (Deines et al., 2019; Glose et al., 2022) and in parts of Texas (Mrad et al., 2020) and which are proposed as part of California's Sustainable Groundwater Management Act (SGMA) (Berbel and Esteban, 2019; Lubell et al., 2020), could also help to improve the efficiency of irrigation and, hence, the ability to more effectively meet crop water demands with lower well capacities. Such interventions may reduce - but not eliminate - increases in drought risk caused by lower or declining aquifer thickness, and could also help to mitigate against potential yield losses caused by over-use of water in wetter years via raising awareness of efficiency saving potentials (Foster et al., 2019) and the importance of conservation (Marston et al., 2022). However, risks also exist that irrigation efficiency improvements will increase net extraction, ultimately causing a rebound effect that exacerbates future aquifer depletion and drought risks (Grafton et al., 2018; Perez-Blanco et al., 2021)."}

\textcolor{blue}{New references added:}

\begin{itemize}
\item
  \textcolor{blue}{Glose, T. J., Zipper, S., Hyndman, D. W., Kendall, A. D., Deines, J. M., and Butler, J. J. (2022). Quantifying the impact of lagged hydrological responses on the effectiveness of groundwater conservation. Water Resources Research, 58(7), e2022WR032295.}
\item
  \textcolor{blue}{Berbel, J., and Esteban, E. (2019). Droughts as a catalyst for water policy change. Analysis of Spain, Australia (MDB), and California. Global Environmental Change, 58, 101969.}
\item
  \textcolor{blue}{Lubell, M., Blomquist, W., and Beutler, L. (2020). Sustainable groundwater management in California: A grand experiment in environmental governance. Society and Natural Resources, 33(12), 1447-1467.}
\item
  \textcolor{blue}{Marston, L. T., Zipper, S., Smith, S. M., Allen, J. J., Butler, J. J., Gautam, S., and David, J. Y. (2022). The importance of fit in groundwater self-governance. Environmental Research Letters, 17(11), 111001.}
\item 
  \textcolor{blue}{Pérez-Blanco, C. D., Loch, A., Ward, F., Perry, C., and Adamson, D. (2021). Agricultural water saving through technologies: a zombie idea. Environmental Research Letters, 16(11), 114032.}
\item 
  \textcolor{blue}{Grafton, R.Q., Williams, J., Perry, C.J., Molle, F., Ringler, C., Steduto, P., Udall, B., Wheeler, S.A., Wang, Y., Garrick, D. and Allen, R.G. (2018). The paradox of irrigation efficiency. Science, 361(6404), 748-750.}
\item 
  \textcolor{blue}{Mrad, A., Katul, G.G., Levia, D.F., Guswa, A.J., Boyer, E.W., Bruen, M., Carlyle-Moses, D.E., Coyte, R., Creed, I.F., Van De Giesen, N. and Grasso, D. (2020). Peak grain forecasts for the US High Plains amid withering waters. Proceedings of the National Academy of Sciences, 117(42), 26145-26150.}
\end{itemize}

\hypertarget{minor-points}{%
\subsection{Minor points}\label{minor-points}}

\begin{itemize}
\tightlist
\item
  Section 1, 3rd Paragraph -- ``Two aquifer pathways exist through aquifer depletion will.'' – confusing sentence – please replace ``through'' with ``by which''
\end{itemize}

\textcolor{blue}{We have corrected the text in line with the reviewer's comment.}

\begin{itemize}
\tightlist
\item
  Section 1, 4th Paragraph --``we generate empirical evidence'' seems awkward to me. How about replacing with ``we present empirical evidence''? Next sentence – ``Our analysis advances on previous $\dots$'' reads better as ``Our analysis extends previous $\dots$'' There are many issues such as these and the previous comment. Please address such issues so that the manuscript is an easier read.
\end{itemize}

\textcolor{blue}{We completely agree. Both of the suggested changes have been made.}

\begin{itemize}
\tightlist
\item
  Section 2.1 title -- I would delete ``marginally'\,' as it is not
  needed and some will find it confusing.
\end{itemize}

\textcolor{blue}{We agree and have removed this term for clarity.}

\begin{itemize}
\tightlist
\item
  Figure 2 - Why do the plots stop at a thickness of about 12 m when you
  had earlier defined the minimum thickness as 9 m and mention 10 m in
  the text when discussing the relationships?
\end{itemize}

\textcolor{blue}{Thank you for catching this typo. The correct value is 9 m and the text has been modified to reflect this and ensure consistency. Plots have also been updated to finish at an aquifer thickness of 9 m where relevant.}

\begin{itemize}
\tightlist
\item
  Section 2.3, First Paragraph -- please clarify that ``average per-area
  production'\,' is a combination of irrigated and rainfed yields. That
  would make the manuscript easier to follow.
\end{itemize}

\textcolor{blue}{We have modified the text in Section 2.3, first paragraph to now more clearly state: "weighted-average of irrigated and rainfed yields, where weights are production area shares".}

\begin{itemize}
\tightlist
\item
  Figure 5 -- I assume that the left plot is corn and the right soybean,
  but please clarify.
\end{itemize}

\textcolor{blue}{Thanks for noticing this. We have added this detail to the figure and its caption.}

### 2023-08-30 Writing/NatureWater/1stRoundRevision/replies/cover_letter.tex
added filed from overleaf for syncing

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Undesirable Weather},
  pdfauthor={Taro Mieno; Timothy Foster; Shunkei Kakimoto; Nicholas Brozovic},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

Dear Editor,\\

Thank you for considering this manuscript for publication at Nature Water. In this study, we demonstrate the importance of aquifer conservation for ensuring agricultural resilience to current and future drought events. Using the High Plains Aquifer in the central United States as a case study, we show that declines in aquifer saturated thickness significantly reduces the overall productivity of both corn and soybean during drought events. Negative impacts on agricultural productivity during droughts occur both through reductions in per-area crop yields, and reductions in expected irrigated areas as farmers are forced to transition land to rainfed production as a result of water supply constraints caused by aquifer depletion. Our findings demonstrate that unsustainable groundwater use is likely to substantially increase vulnerability of agriculture to drought in many parts of the world, and highlight the need for greater focus on proactive conservation of groundwater stocks to support adaptation of farms, agricultural supply chains, and rural economies to climate change.\\

Our work is significant because we quantify the impact of a decline in saturated thickness of groundwater on the resiliency of agricultural production, which has rarely been quantified in the literature before. We confirm quantitatively the importance of keeping relatively high saturated thickness for sustainable high-productivity agricultural production. While our results are for the US High Plains, the insights we generated here are likely to hold elsewhere.\\

This article does not have any realated manuscripts that are under consideration. However, its pre-print has been published at EarthArXiv (\url{https://eartharxiv.org/repository/view/5126/}). We have not had any prior discussions with a Nature Water editor about this work. \\

Here is the list of potential reviewers that we may recommend.

\begin{itemize}
    \item James Butler (\url{https://geo.ku.edu/people/james-k-butler-jr})
    \item Marc Muller (\url{https://engineering.nd.edu/faculty/marc-francois-muller/})
    \item Sam Zipper (\url{https://www.samzipper.com/authors/samuel-c.-zipper/})
    \item Christian Klassert (\url{https://www.ufz.de/index.php?en=38767})
    \item Jen Burney (\url{http://www.jaburney.net/})
\end{itemize}

We do not have any reviewers that we would like to avoid.

\end{document}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Aquifer depletion exacerbates agricultural drought losses in the US High Plains},
  pdfauthor={Taro Mieno; Timothy Foster; Shunkei Kakimoto; Nicholas Brozovic},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

Dear Dr. Sjostrand,\\

Thank you for the opportunity to revise and resubmit our manuscript. We have addressed all the comments provided by the reviewers in our new manuscript, and have included detailed responses to each point alongside our revised manuscript.

Specifically, in our revised manuscript, we have made the following major changes based on the reviewers' comments, alongside several minor adjustments to wording and to correct spelling and grammatical errors:

\begin{itemize}
    \item Updated the main results section to include additional variables (e.g., soil characteristics, state fixed effects) to strengthen how we account for unobservable factors in our regression analyses.
    \item Added several new figures (Figure 2 and Figure 5) to more clearly highlight the non-linearity and statistical significance of declining saturated thickness on per-area crop yields (Figure 2) and overall crop productivity (Figure 5), which, combined with revised discussion of our results, we believe enable readers to gain deeper understanding of the significance of our findings. 
    \item Conduct rigorous statistical testing of our main findings, in particular to highlight the specific climate and aquifer conditions at which aquifer depletion has statistically robust and significant impacts on drought impacts and risks, including how these thresholds vary between corn and soybean.
    \item Strengthened the descriptions on the underlying theoretical and empirical evidence for causal links between crop yield, share of irrigated acres, and aquifer thickness, which are one of the main empirical findings from our paper.
    \item Added a series of four appendices to discuss in more details our methods, robustness checks, limitations imposed by underlying data. 
\end{itemize}

We believe that these changes have substantially strengthen both the methodological approaches in our manuscript, as well the implications of our findings for groundwater conservation and management.

Thank you once for again for consideration of our manuscript, and we would be happy to provide any further information that you may require.

Sincerely,

Taro Mieno (on behalf of the authors)
\end{document}
